Time Series Analysis ARIMA basic model

https://medium.com/@kfoofw/seasonal-lags-sarima-model-fa671a858729
```{r, message = FALSE}

library(ggplot2)
library(tidyquant)
library(tidyverse)
library(dplyr)
library(data.table)
library(corrplot)
library(gridExtra)
#time series libraries
library(forecast)
library(tseries)
library(TSA)
library(sarima)
library(astsa)

library(tibble)
library(TTR)
library(xts)
library(dygraphs)

#save images
library(htmlwidgets)
library(webshot)

library(assertthat)


```

# Import time series data
```{r}
cdc_dec20_ts = read.csv('cdc_dec20_ts.csv') %>% 
  transmute(Date = as.Date(X2020.01.01),
            Values = X2) %>% 
  mutate(Group = "Given")

cdc_xts = xts(cdc_dec20_ts$Values, order.by = cdc_dec20_ts$Date)

cdc_may21_ts = read.csv('cdc_may21_ts.csv') %>% 
  transmute(Date = as.Date(X2020.01.01),
            Values = X15) %>% 
  mutate(Group = "Given")

cdc_xts_may = xts(cdc_may21_ts$Values, order.by = cdc_may21_ts$Date)


cdc_allmay.ts = read.csv('cdc_allmay_positive_cases_ts.csv') %>% 
  transmute(Date = as.Date(X2020.01.01), Values = X45) %>% 
  mutate(Group = "Given")

cdc_xts_allmay = xts(cdc_allmay.ts$Values, order.by = cdc_allmay.ts$Date)
```

# Plot
```{r}
dygraph(cdc_xts, xlab = "Time", ylab = "Cases", main = "Time Series")

```

```{r}
dygraph(cdc_xts_may, xlab = "Time", ylab = "Cases", main = "Time Series")
```

```{r}
######################## Include in report
#https://github.com/rstudio/dygraphs/issues/71
# save html to png

allmay_tsplot = dygraph(cdc_xts_allmay, xlab = "Time", ylab = "Cases", main = "Partial Time Series of Positive COVID Cases")

saveWidget(allmay_tsplot, "allmay_dygraph.html", selfcontained = FALSE)
width<- 1080
height <- 610
webshot("allmay_dygraph.html", file = "allmay_dygraph.png",
        cliprect = c(10,30,width+50,height+50)
        ,vwidth = width, vheight = height )



```




# Stationarity Testing
```{r}
# test for stationarity
# Augmented Dickey-Fuller test
adf.test(cdc_xts, alternative = "stationary", k = 0)
```

```{r}
# test for stationarity
# Augmented Dickey-Fuller test
adf.test(cdc_xts_may, alternative = "stationary", k = 0)
```

```{r}
# INCLUDE THIS

# test for stationarity
# Augmented Dickey-Fuller test
adf.test(cdc_xts_allmay, alternative = "stationary", k = 0)
```

```{r}
cdc_xts_diff1 = diff(cdc_xts)
plot.xts(cdc_xts_diff1, col = "dark green")
```

```{r}
cdc_xts_may_diff1 = diff(cdc_xts_may)
plot.xts(cdc_xts_may_diff1, col = "dark green")
```
```{r}
png(file = "allmay_plotxts.png", width = 1080, height = 610)
plot.xts(diff(cdc_xts_allmay), col = "dark green")
dev.off()
```


```{r}
adf.test(na.omit(cdc_xts_diff1), alternative = "stationary", k = 0)
```

```{r}
adf.test(na.omit(cdc_xts_may_diff1), alternative = "stationary", k = 0)
```

```{r}
adf.test(diff(cdc_xts_allmay) %>% na.omit(), alternative = "stationary", k = 0)
```

```{r}
findfrequency(cdc_xts)
findfrequency(cdc_xts_diff1)

findfrequency(cdc_xts_may)
findfrequency(cdc_xts_may_diff1)

findfrequency(cdc_xts_allmay)
findfrequency(diff(cdc_xts_allmay))
```

# ACF/PACF plot: Nov xts
```{r}
#Acf gradually declines.
#Acf(cdc_xts_diff1, lag.max = 60)
acf2(diff(cdc_xts),8)
cdc_nov_acfObj = acf2(diff(diff(cdc_xts),7),8)
```

# ACF: May xts
```{r}
#Acf gradually declines.
#Acf(cdc_xts_may_diff1, lag.max = 60)

acf2(diff(cdc_xts_may),12)

#We're using this one 
cdc_may_acfObj = acf2(diff(diff(cdc_xts_may),7))
```

#ACF/PACF: all May xts
```{r}
#Acf gradually declines.
#Acf(cdc_xts_may_diff1, lag.max = 60)

acf2(cdc_xts_allmay)
acf2(diff(cdc_xts_allmay))
acf2(diff(cdc_xts_allmay),12)
#We're using this one 

cdc_allmay_acfObj = acf2(diff(diff(cdc_xts_allmay),7))
```

#https://towardsdatascience.com/identifying-ar-and-ma-terms-using-acf-and-pacf-plots-in-time-series-forecasting-ccb9fd073db8

The plots suggests, we use an autoregressive model with lag = 7. If ACF plot shows sharp drop off and PACF plot shows gradual decline, we go with MA model. If ACF plot shows gradual decline and PACF shows sharp drop off, we go with AR model. If both shows gradual decline, we use ARMA model.


```{r}
acf2(diff(diff(cdc_xts),7),12)
sarima(cdc_xts, p = 0, d = 1, q = 2, P = 2, D = 1, Q = 2, 7)
sarima(cdc_xts, p = 2, d = 1, q = 2, P = 2, D = 1, Q = 2, 7)
```

#Not using
```{r}
acf2(diff(diff(cdc_xts_may),7),12)
sarima(cdc_xts_may, p = 0, d = 1, q = 2, P = 2, D = 1, Q = 2, 7)
sarima(cdc_xts_may, p = 2, d = 1, q = 2, P = 2, D = 1, Q = 2, 7)
```
# Use this one
```{r}
####################### Include in report
png(file = "allmay_acf2.png", width = 1080, height = 610)
acf2(diff(cdc_xts_allmay))
dev.off()
png(file = "allmay_acf2_lag7.png", width = 1080, height = 610)
acf2(diff(diff(cdc_xts_allmay),7))
dev.off()
#sarima(cdc_xts_allmay, p = 0, d = 1, q = 2, P = 2, D = 1, Q = 2, 7)

####################### Include in report
png(file = "allmay_sarima.png", width = 1080, height = 610)
sarima(cdc_xts_allmay, p = 2, d = 1, q = 2, P = 2, D = 1, Q = 2, 7)
dev.off()
```

```{r}
days = 60

#sarima forecast for 5 days with parameters (2, 1, 2) (2, 1, 2)_7
png(file = "allmay_sarima_60days_2021.png", width = 1080, height = 610)
sarima.allmay = sarima.for(cdc_xts_allmay, days, p = 2, d = 1, q = 2, P = 2, D = 1, Q = 2, 7) 
dev.off()

sarima.allmay_pred = sarima.allmay$pred %>% map(~ max(0,.x)) %>% unlist()


allmay_pred_df = data.frame(Date = seq(from = tail(cdc_may21_ts$Date,1)+1, by = 1, length.out = days),
                         Values = sarima.allmay_pred,
                         Group = "Forecasted")

allmay_full = rbind(cdc_allmay.ts, allmay_pred_df)

ggplot(data = allmay_full, aes(x = Date, y = Values, color = Group, group = 1)) + geom_line() + coord_x_date(xlim = c("2021-03-01","2021-15-05"), ylim = (c(-10,20000))) + ggtitle("Time Series 60 day forecast (adjusted for negative values)")

ggsave("allmay_ggplot_60days_2021.png")
```

#Comparison with the ground truth, predicting the winter surge
```{r}
days = 60
apr_oct2020 = cdc_allmay.ts %>% filter(Date <= "2020-10-31")

apr_oct2020.xts = xts(apr_oct2020$Values, order.by = apr_oct2020$Date)

```

```{r}
acf2(diff(apr_oct2020.xts))
acf2(diff(diff(apr_oct2020.xts),7))
```

```{r}
days = 60
png(file = "allmay_sarima_60days_2020.png", width = 1080, height = 610)
sarima.aproct = sarima.for(apr_oct2020.xts, days, p = 2, d = 1, q = 2, P = 2, D = 1, Q = 2, 7)
dev.off()

aproct_pred_df = data.frame(Date = seq(from = tail(apr_oct2020$Date,1)+1, by = 1, length.out = days),
                         Values = sarima.aproct$pred,
                         Group = "Forecasted")

aproct_full = rbind(apr_oct2020, aproct_pred_df)

aproct_predplot = ggplot(data = aproct_full, aes(x = Date, y = Values, color = Group, group = 1)) + geom_line() + coord_x_date(ylim = (c(-10,80000))) + ggtitle("Time Series")

aproct_actplot = ggplot(data = cdc_allmay.ts, aes(x = Date, y = Values)) + geom_line() + coord_x_date(xlim = c("2020-01-01","2021-01-01"),ylim = (c(-10,80000))) + ggtitle("Time Series")

aproct_predplot
ggsave("allmay_ggplot_60days.png")
aproct_actplot
```

```{r}
actual_toforecast_may = cdc_allmay.ts %>% 
  filter(Date > tail(apr_oct2020$Date) + 1) %>% 
  filter(Date < tail(apr_oct2020$Date) + days)

ggplot(data = aproct_full, aes(x = Date, y = Values, color = Group, group = 1)) + geom_line()+ geom_line(data = actual_toforecast_may, aes(x = Date, y = Values, color = Group, group = 1)) + coord_x_date(xlim = c("2020-01-01","2021-01-01"),ylim = (c(-10,80000))) + ggtitle("Time Series 60 day forecast (comparison to actual)")

ggsave("allmay_ggplot_60days_2020.png")
```




